# Glossary {.unnumbered}


## Boxplot

A boxplot is a visual method of showing a distribution of values, it is popular among statisticians and is a useful way to explore data. @fig-boxexample shows the essential outline of the boxplot and the text following explains their relevance:

```{r}
#| include: false
library(datasets)
library(timetk)
library(tidyverse)

data(Nile)

NileData <- timetk::tk_tbl(Nile)

#Add outlier
outlier <- tibble(index = 1870, value = 1800)
NileData <- rbind(NileData, outlier) %>% 
  arrange(index)

```

```{r}
#| label: fig-boxexample
#| fig-cap: "Example boxplot with whiskers and outlier"
#| warning: false

ggplot(NileData, aes(x = "", y = value)) +
        geom_boxplot(fill = "grey85") +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.title.x = element_blank(),
              panel.background = element_blank(),
              axis.title.y = element_blank(),
              axis.text.x = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks = element_blank(),
              panel.border = element_rect(colour = "black", fill=NA, size=2))

```


Each boxplot consists of a box that stretches from the 25th percentile ($Q_1$) of the distribution to the 75th percentile ($Q_3$), a distance known as the interquartile range ($IQR$). In the middle of the box is a line that displays the median, i.e. 50th percentile, of the distribution. These three lines give you a sense of the spread of the distribution and whether or not the distribution is symmetric about the median or skewed to one side. The _dots_ at the end of the boxplot represent outliers There are a number of different rules for determining if a point is an outlier, but the method that R [@R-base] and ggplot [@R-ggplot2] use is the _1.5 rule_. 

If a data point is:

* less than $Q_1 - 1.5 * IQR$  
* greater than $Q_3+ 1.5 * IQR$  

then that point is classed as an _outlier_. IQR is Inter Quartile Range 

The whiskers are defined as:

* upper whisker = $min(max(x), Q_3 + 1.5 * IQR)$

* lower whisker = $max(min(x), Q_1 - 1.5 * IQR)$

where $IQR = Q_3 - Q_1$, the box length. So the upper whisker is located at the smaller of the maximum x value and $Q_3 + 1.5 * IQR$, whereas the lower whisker is located at the larger of the smallest x value and $Q_1 - 1.5 * IQR$.

Additional information about boxplots can be found <https://en.wikipedia.org/wiki/Box_plot>
