# Example Data


```{r}
#| include: false

library(knitr)
library(tidyverse)
library(imagic)
library(broman)
library(gt)
library(gtExtras)
library(glue)
library(leaflet)
```

## The `WwTW_hydro` Data

```{r}
#| label: fig-hydrodatamap
#| fig-cap: "Map of the `WwTW_hydro` Data"


WwTW_hydro %>%
    mutate(Popup_text = glue("{WWTP_NAME} - population served {add_commas(POP_SERVED)}")) %>% 
    leaflet() %>% 
    addTiles() %>% 
    addCircleMarkers(
      lng = ~LON_WWTP, lat = ~LAT_WWTP,
      radius = 7,
      #color = ~pal(Species),
      stroke = TRUE, fillOpacity = 0.7,
      popup = ~Popup_text, label = ~Popup_text
    )

```

@fig-hydrodatamap shows the spacial extent of the `WwTW_hydro` data in the `imagic` package. The data has information on `r add_commas(nrow(WwTW_hydro))` Wastewater Treatment Works (WwTW); hovering on the map will _pop up_ the names of the works and population served. The data is a subset of South Africa from the global wastewater treatment plant database used in the paper published in 2022[^1] 

[^1]:Ehalt Macedo, H., Lehner, B., Nicell, J. A., Grill, G., Li, J., Limtong, A., Shakya, R.: Distribution and characteristics of wastewater treatment plants within the global river network. Earth System Science Data. 2022.

Other information in the  `WwTW_hydro` data:

* ORG_ID		ID from national/regional dataset (see reference paper for more information)
* WWTP_NAME	Name of the WwTW from national/regional dataset (empty if not reported)
* COUNTRY		Country in which WwTW is located
* CNTRY_ISO	Country ISO
* LAT_WWTP	Latitude of reported WwTW location
* LON_WWTP	Longitude of reported WwTW location
* QUAL_LOC	Quality indicator related to reported WwTW location (see SI of reference paper for more information): 1 = high (tests indicated >80% of reported WwTW locations in country/region to be accurate); 2 = medium (tests indicated between 50% and 80% of reported WwTW locations in country/region to be accurate); 3 = low (tests indicated <50% of reported WwTW locations in country/region to be accurate); 4 = Quality of WwTW locations in country/region not analysed
* LAT_OUT		Latitude of the estimated outfall location (see reference paper for more information)
* LON_OUT		Longitude of the estimated outfall location (see reference paper for more information)
* STATUS		Status of the WwTW from national/regional dataset: Closed, Construction Completed, Decommissioned, Non-Operational, Operational, Projected, Proposed, Under Construction, Not Reported (assumed operational)
* POP_SERVED	Population served by the WwTW
* QUAL_POP	Quality indicator related to the attribute "population served" (see reference paper for more information): 1 = Reported as ‘population served’ by national/regional dataset; 2 = Reported as ‘population equivalent’ by national/regional dataset; 3 = Estimated (with wastewater discharge available); 4 = Estimated (without wastewater discharge available)
* WASTE_DIS	Treated wastewater discharged by the WwTW in m3 d-1
* QUAL_WASTE	Quality indicator related to the attribute "Treated wastewater discharged" (see reference paper for more information): 1 = Reported as ‘treated’ by national/regional dataset; 2 = Reported as ‘design capacity’ by national/regional dataset; 3 = Reported but type not identified; 4 = Estimated
* LEVEL		Level of treatment of the WwTW: Primary, Secondary, Advanced
* QUAL_LEVEL	Quality indicator related to the attribute "level of treatment" (see reference paper for more information): 1 = Reported by national/regional dataset; 2 = Estimated
* DF		Estimated dilution factor (empty if estimated outfall location is the ocean or large lake; see reference paper for more information)
* HYRIV_ID	ID of associated river reach in RiverATLAS at estimated outfall location (link to HydroATLAS database; empty if estimated outfall location is the ocean or an endorheic sink)
* RIVER_DIS	Estimated river discharge at the WwTW outfall location in m3 s-1 (derived from HydroATLAS database; empty if estimated outfall location is the ocean)
* COAST_10KM	1 = Estimated outfall location within 10 km of the ocean or a large lake (surface area larger than 500 km2); 0 = Estimated outfall location further than 10 km of the ocean or a large lake (surface area larger than 500 km2)
* COAST_50KM	1 = Estimated outfall location within 50 km of the ocean or a large lake (surface area larger than 500 km2); 0 = Estimated outfall location further than 50 km of the ocean or a large lake (surface area larger than 500 km2)
* DESIGN_CAP	Design capacity of WwTW as reported in national/regional dataset (empty if not reported)
* QUAL_CAP	Quality indicator related to the attribute "design capacity": 1 = Reported as design capacity in m3 d-1; 2 = Reported as design capacity in 'population equivalent'; 3 = Not reported

## The `wtw_data` Water Treament Quality Data

@tbl-imagicdata shows the structure of the `wtw_data` in the `imagic` package. The data has `r add_commas(nrow(wtw_data))` rows.


```{r}
#| label: tbl-imagicdata
#| tbl-cap: "Data From wtw_data in the imagic Package"

wtw_data %>% 
  sample_n(10) %>% 
  gt() %>% 
  tab_style(
    style = list(
      cell_fill(color = "black"),
      cell_text(color = "white", weight = "bold")
      ),
    locations = cells_column_labels(
      columns = c(everything())
    )
  ) 


```



```{r}
#| label: tbl-datacount
#| tbl-cap: "Water Treatment Works (WTW) and the Number of Determinands in wtw_data"
#| warning: false


wtw_data %>% 
  mutate(Description = stringr::str_replace(Description, " WTW", "")) %>% 
  group_by(Description, Determinand) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(n = add_commas(n)) %>% 
  pivot_wider(names_from = Description, values_from = n) %>% 
  gt() %>% 
  tab_style(
    style = list(
      cell_fill(color = "black"),
      cell_text(color = "white", weight = "bold")
      ),
    locations = cells_column_labels(
      columns = c(everything())
    )
  ) %>% 
   cols_align(
    align = "right",
    columns = 2:10
  )
  # gt_hulk_col_numeric(2:10, trim = TRUE) 
  #gt_color_rows(2:10, palette = "ggsci::blue_material") 



```