# Timeline Graphs {#ChapSGraphs}


```{r}
#| include=FALSE

library(tidyverse)
library(RColorBrewer)
library(ggthemes)
library(scales)
library(knitr)
library(scales)
library(imagic)
library(glue)

```



This chapter illustrates the range of graphs that can help describe the hidden relationships in the drinking water quality data over time.  

## Plots of Determinands against Time 

### Determinand on a Linear Scale 

An example determinand time line using a linear scale is shown in @fig-timeline1 for temperature at _Mirkwood WTW_. Temperatures would normally be plotted using a linear scale; other determinands that would _generally_ be plotted on a linear scale are:

* pH  
* Free Chlorine 
* Total Chlorine
* Colour  
* Conductivity  
* Turbidity

```{r}
#| label: fig-timeline1
#| fig-cap: "Example Time Line Plot on a Linear Scale"

# filter the wtw_data with the filter function and the dplyr package
mirkwood_temp_data <- wtw_data %>%
  filter(Description == "Mirkwood WTW", Determinand == "Temperature")

ggplot(mirkwood_temp_data, aes(x = Sample_Date, y = Result)) +
  geom_point(size = 1, alpha = 0.9, colour = "steelblue") +
  labs(y = "Temperature (\u00B0C)") +
  scale_x_date(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  facet_grid(Description ~ .) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 270, hjust = 1)) +
  theme(axis.title.x = element_blank())

```

### Determinand on a Log Scale  

An example determinand time line using a log scale is shown in @fig-timeline2 for _Colony Count 21°C_ at _Mirkwood WTW_. By convention, and clarity, microbiological determinands are displayed on a log10 scale.


```{r}
#| label: fig-timeline2
#| fig-cap: "Example Time Line Plot on a Log10 Scale"
#| warning: false

# filter the wtw_data with the filter function and the dplyr package
mirkwood_CC_data <- wtw_data %>%
  filter(Description == "Mirkwood WTW", Determinand == "Colony Count 21")

ggplot(mirkwood_CC_data, aes(x = Sample_Date, y = Result)) +
  geom_point(size = 1, alpha = 0.9, colour = "yellow4") +
  labs(y = "Colony Count 21\u00B0C (mL)") +
  scale_x_date(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  facet_grid(Description ~ .) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 270, hjust = 1)) +
  theme(axis.title.x = element_blank()) +
  annotation_logticks(sides = "rl")

```